---
title: "Foundations of Data Science: Exercise 4 - Titanic data visualisation"
author: "James Hooi"
date: "October 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load required libraries
library(ggplot2)
library(dplyr)
# Set working directory
setwd("C:\\Users\\james.hooi\\Documents\\data analytics\\Springboard FDS exercises\\Ex4.0.1 Exploratory Data Viz\\FDS-Ex4-titanic")
# Get data and clean it
titanic.clean <- read.csv(file = paste(getwd(),"titanic_clean.csv",sep="\\"))
titanic <- titanic.clean %>% 
  select(Survived = survived, Pclass = pclass, Sex = sex, Age = age)
titanic <- titanic %>% filter(Sex != "")  
```

## Titanic: exploratory data visualisations

Using cleaned data from Ex 3.1.2, note there are more records in this dataset than the Data Camp version.

###1. Check the structure of the data
```{r task1}
str(titanic)
```

###2. Plot bar graph of passenger survival by class and by sex
```{r task2}
ggplot(titanic, aes(x = Pclass, fill = Sex)) + 
  geom_bar(position = "dodge") +
  labs(x = "Passenger Class", y = "Count", title = "Bar plot: Passenger survival by class, by sex")
```

###3. Plot bar graph of passenger survival by class and by sex, faceted by survival
```{r task3}
ggplot(titanic, aes(x = Pclass, fill = Sex)) + 
  geom_bar(position = "dodge") +
  facet_grid(. ~ Survived) +
  labs(x = "Passenger Class", y = "Count", title = "Bar plot: Passenger survival by class, by sex"
       , subtitle = "LHS: did not survive, RHS: survived")
```

###4. & 5. Define jitter-dodge position for dot plot and draw
- Passenger survival by class, by sex, by age
- Faceted by survival
```{r task5}
posn.jd <- position_jitterdodge(0.5, 0, 0.6)

ggplot(titanic, aes(x = Pclass, y=Age, col = Sex)) + 
  geom_point(position = posn.jd, size=3, alpha=0.5) +
  facet_grid(. ~ Survived) +
  labs(x = "Passenger Class", title = "Dot plot: Passenger survival by class, by sex, by age"
       , subtitle = "LHS: did not survive, RHS: survived")
```

